\bf Soundness Proofs for Iterative Deepening

### 2016-09-06 oopsla-16/proof/proof.tex
oopsla proofs: revise convergence proofs for prosaic issues
Finish oopsla proofs for TRification

%
% the following standard packages may be helpful, but are not required
%
\usepackage{SIunits}            % typset units correctly
\usepackage{courier}            % standard fixed width font
\usepackage[scaled]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
\usepackage{url}                  % format URLs
\usepackage{listings}          % format code
\usepackage{amsthm}          % format code
\usepackage{enumitem}      % adjust spacing in enums
\usepackage[linesnumbered,ruled]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{graphicx}
%\usepackage{setspace}
\usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}   % hyperlinks, including DOIs and URLs in bibliography
% known bug: http://tex.stackexchange.com/questions/1522/pdfendlink-ended-up-in-different-nesting-level-than-pdfstartlink
\newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}   % print a hyperlinked DOI
%\doublespacing

@@

%\title{Soundness Proof for Eliminating Malloc-Recycle Data Races in Stateless Model Checking}
\title{Soundness Proofs for Iterative Deepening}
\authorinfo{Ben Blum}{Carnegie Mellon University}{bblum@cs.cmu.edu}

\maketitle
\begin{abstract}
In our paper we show the effectiveness of combining dynamic data-race detection \cite{eraser,hybriddatarace} with stateless model checking \cite{verisoft,dpor}.
Our approach involves adding new state spaces to explore each time a new data-race candidate is found.
This document presents two proofs concerning the technique.

@@

Second, we prove the soundness of the {\em malloc-recycle} false-positive-elimination tactic we discuss in the paper.
Despite many powerful reduction techniques, the state spaces are still exponentially sized,
so any way to avoid exploring some in advance is obviously beneficial.
We prove that for the special case of {\em malloc-recycle} false positives, it is safe to eliminate these immediately upon discovering them, without bothering to explore their associated state spaces.

	This document is supplemental material to our main paper \cite{quicksand}; we assume the reader is already familiar with our motivation and terminology.
\end{abstract}

@@

\section{Acknowledgements}

%{\em removed for double blind review}
Michael J. ``Sully'' Sullivan, Carlo Angiuli, and Nathaniel Filardo double-checked the proof and provided valuable feedback on notation style and the flow of explanation.
Thanks as well to the anonymous OOPSLA reviewer \#4, whose suggestions helped improve the formality of our definitions.

This work was supported in part by
the member companies of the PDL Consortium (Broadcom,
Citadel, EMC, Facebook, Google, Hewlett-Packard Labs, Hitachi, Intel,
Microsoft Research, MongoDB, NetApp, Oracle, Samsung America, Seagate,
Tintri, Two Sigma, Uber, Veritas and Western Digital)
and by the U.S. Army Research Office under grant number W911NF0910273.

>>>>

%
% the following standard packages may be helpful, but are not required
%
\usepackage{SIunits}            % typset units correctly
\usepackage{courier}            % standard fixed width font
\usepackage[scaled]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
\usepackage{url}                  % format URLs
\usepackage{listings}          % format code
\usepackage{amsthm}          % format code
\usepackage{enumitem}      % adjust spacing in enums
\usepackage[linesnumbered,ruled]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{cmu-techreport}
\usepackage{flushend}
%\usepackage{setspace}
\usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}   % hyperlinks, including DOIs and URLs in bibliography
% known bug: http://tex.stackexchange.com/questions/1522/pdfendlink-ended-up-in-different-nesting-level-than-pdfstartlink
\newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}   % print a hyperlinked DOI
%\doublespacing

@@

%\title{Soundness Proof for Eliminating Malloc-Recycle Data Races in Stateless Model Checking}
\title{\bf Soundness Proofs for Iterative Deepening}
%\authorinfo{Ben Blum}{Carnegie Mellon University}{bblum@cs.cmu.edu}
\author{Ben Blum}

% tech report stuff
\trnumber{CMU-PDL-16-103}
\keywords{concurrency, stateless model checking, data race analysis, verification}

\abstract{
	\renewcommand{\baselinestretch}{1.0}
	\small
	\em
	%We have developed the
	The Iterative Deepening algorithm allows stateless model checkers to adjust preemption points on-the-fly.
	%incorporate dynamic data-race analysis during testing.
	%The algorithm
	It uses dynamic data-race detection to avoid necessarily preempting on every shared memory access,
	and ignores false-positive data race candidates arising from certain heap allocation patterns.
	%When an Iterative Deepening test reaches completion,
	An Iterative Deepening test that reaches completion soundly verifies all possible thread interleavings of that test.
}

\support{\small{
	{\bf Acknowledgements:}
	Michael J. ``Sully'' Sullivan, Carlo Angiuli, and Nathaniel Filardo double-checked the proof and provided valuable feedback on notation style and the flow of explanation.
Thanks as well to the anonymous OOPSLA reviewer \#4, whose suggestions helped improve the formality of our definitions.
%This work was supported in part by
We thank
the member companies of the PDL Consortium (Broadcom,
Citadel, EMC, Facebook, Google, Hewlett-Packard Labs, Hitachi, Intel,
Microsoft Research, MongoDB, NetApp, Oracle, Samsung America, Seagate,
Tintri, Two Sigma, Uber, Veritas and Western Digital)
and by the U.S. Army Research Office under grant number W911NF0910273.
}}

\maketitle

\section{Introduction}
In our OOPSLA paper \cite{quicksand} we introduced Iterative Deepening,
an algorithm which combines
stateless model checking \cite{verisoft,dpor}
and
data-race analysis \cite{eraser,hybriddatarace}.
It dynamically adds new state spaces to test each time a data-race candidate is found.
We present here two proofs concerning the technique.

@@

Second, we prove the soundness of the {\em malloc-recycle} false-positive-elimination tactic discussed in the paper.
Despite powerful reduction techniques, each state space is still exponentially sized,
%so any way to avoid exploring some in advance is obviously beneficial.
so we wish to avoid exploring any we could prove to be redundant.
We prove that it is safe to eliminate malloc-recycle data race candidates immediately, %upon discovering them,
without bothering to explore their associated state spaces.

This tech report is supplemental material to our main paper; we assume the reader is already familiar with our motivation and terminology.

@@

### 2016-09-06 oopsla-16/proof/recycle.tex
Finish oopsla proofs for TRification

\begin{figure}[t]
	\small
\begin{tabular}{rll}
	& {\bf Thread 1} & {\bf Thread 2} \\
	1 & & \texttt{z = malloc(42);} \\
	%2 & & \texttt{\hilight{commentblue}{// TODO bounds check??}} \\
	2 & & \texttt{x2 = \&z[50];} \\
	3 & & \texttt{y~=~\hilight{olivegreen}{malloc}(sizeof *y);} \\
	4 & & \texttt{\hilight{brickred}{x2->foo = ...;}} \\
	5 & \texttt{x1 = malloc(...);} & \\
	6 & \texttt{\hilight{brickred}{x1->foo = ...;}} & \\
	7 & \texttt{\hilight{olivegreen}{free}(x1);} \\
\end{tabular}
\caption{Without a PP between lines 4 and 5 of Figure~\ref{fig:overrun}, this is the only alternate interleaving DPOR would explore. The mallocs have been reordered and may no longer collide, which wrongly appears to be a false positive.}
\label{fig:overrun-notenough}
\end{figure}

\begin{figure}[t]
	\small
\begin{tabular}{rll}
	& {\bf Thread 1} & {\bf Thread 2} \\
	1 & & \texttt{z = malloc(42);} \\
	%2 & & \texttt{\hilight{commentblue}{// TODO bounds check??}} \\
	2 & & \texttt{x2 = \&z[50];} \\
	3 & \texttt{x1 = malloc(...);} & \\
	4 & & \texttt{y~=~\hilight{olivegreen}{malloc}(sizeof *y);} \\
	5 & & \texttt{\hilight{brickred}{x2->foo = ...;}} \\
	6 & \texttt{\hilight{brickred}{x1->foo = ...;}} & \\
	7 & \texttt{\hilight{olivegreen}{free}(x1);} \\
\end{tabular}
	\caption{Goal interleaving of Figure~\ref{fig:overrun}. To ensure collision, the sequence of malloc calls producing $a_1$ and $a_2$ must not be disrupted compared to the original interleaving.}
	%, compared to the trace which originally exposed the malloc-recycle race.}
\label{fig:overrun-goal}
\end{figure}

@@

%This spells QED so we are done.

>>>>

@@

%This spells QED so we are done.
\begin{figure}[t]
	\small
\begin{tabular}{rll}
	& {\bf Thread 1} & {\bf Thread 2} \\
	1 & & \texttt{z = malloc(42);} \\
	2 & & \texttt{\hilight{commentblue}{// TODO bounds check??}} \\
	3 & & \texttt{x2 = \&z[50];} \\
	4 & & \texttt{y~=~\hilight{olivegreen}{malloc}(sizeof *y);} \\
	5 & & \texttt{\hilight{brickred}{x2->foo = ...;}} \\
	6 & \texttt{x1 = malloc(...);} & \\
	7 & \texttt{\hilight{brickred}{x1->foo = ...;}} & \\
	8 & \texttt{\hilight{olivegreen}{free}(x1);} \\
\end{tabular}
\caption{Without a PP between lines 4 and 5 of Figure~\ref{fig:overrun}, this is the only alternate interleaving DPOR would explore. The mallocs have been reordered and may no longer collide, which wrongly appears to be a false positive.}
\label{fig:overrun-notenough}
\end{figure}

\begin{figure}[t]
	\small
\begin{tabular}{rll}
	& {\bf Thread 1} & {\bf Thread 2} \\
	1 & & \texttt{z = malloc(42);} \\
	2 & & \texttt{\hilight{commentblue}{// TODO bounds check??}} \\
	3 & & \texttt{x2 = \&z[50];} \\
	4 & \texttt{x1 = malloc(...);} & \\
	5 & & \texttt{y~=~\hilight{olivegreen}{malloc}(sizeof *y);} \\
	6 & & \texttt{\hilight{brickred}{x2->foo = ...;}} \\
	7 & \texttt{\hilight{brickred}{x1->foo = ...;}} & \\
	8 & \texttt{\hilight{olivegreen}{free}(x1);} \\
\end{tabular}
	\caption{Goal interleaving of Figure~\ref{fig:overrun}. To ensure collision, the sequence of malloc calls producing $a_1$ and $a_2$ must not be disrupted compared to the original interleaving.}
	%, compared to the trace which originally exposed the malloc-recycle race.}
\label{fig:overrun-goal}
\end{figure}
